Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling

### 2021-04-07 manuscript/manuscript.Rmd
work on manuscript

\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}

where the regression parameters (weights) for each layer $k \in [1;K+1]$ are represented by the matrices $a_k$ and the 
intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear 
regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the hyperbolic tangent 
(tanh) function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') 
as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ 
and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of 
the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation).
Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

\begin{equation} 
\label{embed_map}
  z_\mathcal{P}^j : \mathcal{P}^j \to \mathbb{R}^{q_\mathcal{P}^j }, \quad \quad p^j \mapsto  z_\mathcal{P}^j(p),
\end{equation}

where $q_\mathcal{P}^j$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P}^j(.)$ is 
an implicit function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{embed_map} 
states that an embedding maps a level of a categorical variable to a numerical vector. This function is left implicit, 
meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to 
specify exactly how the embeddings can be derived from the input data. In Table \@ref(tab:embed) we show an example of two 
dimensional embeddings for the state variables, where these have been generated randomly.

@@

```{=latex}
\begin{equation}
A: \mathbb{R}^{(q \times d) \times (q \times d)} \to \mathbb{R}^{q \times q}, ~~
(Q, K) ~\mapsto ~ A = \frac{Q~K^T}{\sqrt{d}},
\label{eq:attention_scores}
\end{equation}
```

and where the division by $\sqrt{d}$ is done element-wise. The attention scores $A$ are then normalized, or made to sum to unity, 
across the columns of $A$ using a so-called softmax function i.e.

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is 
made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While 
the 
complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to 
data from 2000 to August 2019 and further subsample 100K data points to make experiments feasible on a CPU.
The dataset includes a rich variety of variables, from occupancy type to flood zone. For our 
models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

[^9]: https://www.fema.gov/about/reports-and-data/openfema

@@

-   Target variable: Amount paid on building claim
-   Predictors:
    - $\log$ `building_insurance_coverage`
    - `basement_enclosure_type`, 
    - `number_of_floors_in_the_insured_building`
    - *prefix* of `flood_zone`
    - `primary_residence`
-   Link function: $\log$
-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to 
match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take 
the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^8]. For example, `"A01"`, 
`"A02"`, and so on are recoded as simply `"A"`. A log link together with the gamma distribution is a standard choice 
for severity modeling, which provides a multiplicative structure where the response is positive. Following the notation
in \@ref(catdata), the number of parameters in the GLM is

\begin{equation}
1 + |J^{\text{num}}| + \left(\sum_{j \in J^{\text{cat}}} n_{\mathcal{P}}^j\right) - |J^{\text{cat}}|:
(\#eq:glm1)
\end{equation}

one for the intercept, one for each numeric variable, and the numbers of levels minus one (due to dummy coding) for 
each categorical variable.

[^8]: https://www.floodsmart.gov/flood-map-zone/find-yours

@@

Table: (\#tab:relativities) Excerpt of relativities for Model 1.

|variable       |label                                       | relativity|
|:--------------|:-------------------------------------------|----------:|
|occupancy_type |Single family residence                     |       1.00|
|occupancy_type |2 to 4 unit residential buidling            |       1.10|
|occupancy_type |Non-residential building                    |       1.34|
|occupancy_type |Residential building with more than 4 units |       1.59|
|flood_zone     |C                                           |       1.00|
|flood_zone     |A                                           |       1.42|
|flood_zone     |X                                           |       1.18|
|flood_zone     |B                                           |       1.65|
|flood_zone     |V                                           |       1.38|
|flood_zone     |D                                           |       1.45|

@@

Table: (\#tab:embeddings-occupancy-type) Learned embeddings for `occupancy_type` in Model 2.

|level                                       | embedding|
|:-------------------------------------------|---------:|
|2 to 4 unit residential buidling            |      1.46|
|Non-residential building                    |     -1.04|
|Residential building with more than 4 units |      0.12|
|Single family residence                     |      0.36|

Table: (\#tab:embeddings-flood-zone) Excerpt of learned embeddings for `flood_zone` in Model 2.

|level | embedding|level | embedding|level | embedding|level | embedding|level | embedding|
|:-----|---------:|:-----|---------:|:-----|---------:|:-----|---------:|:-----|---------:|
|A     |      1.48|A09   |     -0.48|A18   |      0.39|A28   |      1.19|C     |      1.73|
|A00   |     -1.05|A0B   |      1.47|A19   |     -1.06|A30   |     -0.01|D     |     -1.15|
|A01   |     -0.18|A10   |     -0.25|A20   |     -1.01|A99   |     -0.44|V     |     -1.85|
|A02   |     -0.06|A11   |      0.35|A21   |      2.08|AE    |      1.12|V01   |      0.02|
|A03   |     -0.01|A12   |     -0.38|A22   |     -0.15|AH    |      2.19|V02   |     -1.55|
|A04   |      0.05|A13   |      2.81|A23   |     -0.76|AHB   |     -0.68|V03   |     -0.98|
|A05   |      1.30|A14   |      0.58|A24   |     -0.87|AO    |      0.55|V04   |      1.73|
|A06   |     -0.76|A15   |     -0.91|A25   |      0.02|AOB   |      0.08|V05   |     -1.97|
|A07   |     -0.62|A16   |      0.79|A26   |      0.00|AR    |      2.93|V06   |      0.43|
|A08   |     -0.53|A17   |      1.35|A27   |      0.80|B     |      0.33|V07   |      1.91||

@@

\begin{equation}
1 + |J^{\text{num}}| + |J^{\text{cat}}|.
(\#eq:glm2)
\end{equation}

Hence, the model can be written as

@@

\begin{equation}
\mu_i = \beta_0 + \sum_{j \in J^{\text{num}}}\beta_{j} x_{i,j} + \sum_{j \in J^{\text{cat}}} \beta_{j} z_{i,j},
\end{equation}

where $z_{i,j}$ represents the embedding value for $i$th instance of the categorical column $X_j$.

@@

Table: (\#tab:glm2) Fitted coefficients of Model 3.

|term                                     | estimate|
|:----------------------------------------|--------:|
|(Intercept)                              |    3.371|
|total_building_insurance_coverage        |    0.586|
|basement_enclosure_crawlspace_type       |    0.597|
|number_of_floors_in_the_insured_building |   -0.278|
|occupancy_type                           |   -0.075|
|flood_zone                               |   -0.007|
|primary_residence                        |   -0.062|
|community_rating_system_discount         |    0.706|

@@

Table: (\#tab:nn2-occupancy-type) Learned embeddings for `occupancy_type` in Model 4.

|level                                       |    e1|    e2|    e3|
|:-------------------------------------------|-----:|-----:|-----:|
|2 to 4 unit residential buidling            |  0.37| -1.76|  1.36|
|Non-residential building                    |  0.13|  1.47|  1.77|
|Residential building with more than 4 units | -1.12| -0.15| -1.43|
|Single family residence                     |  0.85|  0.16| -0.99|

Table: (\#tab:nn2-flood-zone) Learned embeddings for level `"A00"` of variable `flood_zone` in Model 4.

|dimension |value  |dimension |value  |dimension |value  |dimension |value  |
|:---------|:------|:---------|:------|:---------|:------|:---------|:------|
|e1        |-0.167 |e11       |-0.810 |e21       |-0.300 |e31       |-0.421 |
|e2        |-0.363 |e12       |1.380  |e22       |-0.194 |e32       |0.594  |
|e3        |-0.893 |e13       |-0.180 |e23       |-0.320 |e33       |1.681  |
|e4        |0.130  |e14       |-1.612 |e24       |-0.035 |e34       |1.668  |
|e5        |-1.588 |e15       |-1.327 |e25       |-1.128 |          |       |
|e6        |-0.123 |e16       |-0.965 |e26       |0.168  |          |       |
|e7        |-0.259 |e17       |-1.356 |e27       |-1.022 |          |       |
|e8        |0.324  |e18       |-0.061 |e28       |1.011  |          |       |
|e9        |-1.430 |e19       |-0.120 |e29       |-0.843 |          |       |
|e10       |-1.598 |e20       |-0.488 |e30       |-0.743 |          |       |

@@

\begin{equation}
p_{ij} = \frac{p_{j|i} + p_{i|j}}{2n}.
\end{equation}

On the other hand, in the target space, t-SNE assumes the neighbor selection is driven by a Student 
t-distribution with one
degree of freedom, i.e., a Cauchy distribution. The similarity metric for a pair of points in $\mathcal{Y}$ is 
then defined as the joint probability

\begin{equation}
q_{ij} = \frac{(1 + \norm{y_i - y_j}^2)^{-1}}{\sum_{k \neq l} (1 + \norm{y_k - y_l}^2)^{-1}}.
\end{equation}

With the similarity metrics defined for each space, we can then state the objective for "closeness" for the sets of
points, which is the Kullback–Leibler divergence from Q to P:

\begin{equation}
D_{KL}(P || Q) = \sum_{i \neq j} p_{ij} \log \frac{p_{ij}}{q_{ij}}. (\#eq:tsne-obj)
\end{equation}

We now circle back to Equation \@ref(eq:pji) and discuss $\sigma_i$. The variance of the Gaussian is allowed to vary
depending on where the point is in the space; specifically, it should reflect the density or sparsness of surrounding 
points. To train this parameter, t-SNE introduces a hyperparameter, *perplexity*, which is a scalar value set by the
user, and is defined as

\begin{equation}
Perp(P_i) = 2^{H(P_i)},
\end{equation}

where $H(P_i)$ denotes the entropy of $P_i$,

\begin{equation}
H(P_i) = -\sum_j p_{j|i}\log_2 p_{j|i},
\end{equation}

and $P_i$ is the Gaussian parameterized by $x_i$ and $\sigma_i$. During training, t-SNE looks for $P_i$ that have
perpelxity corresponding to the user input. Finally, the objective as stated in Equation \@ref(eq:tsne-obj) is differentiable, and is trained via gradient descent.

@@

Table: (\#tab:cv-results) Cross validation results.

|Model                                                                         |   RMSE|    MAE|
|:-----------------------------------------------------------------------------|------:|------:|
|Model 1: GLM (gamma/log link)                                                 | 65,138| 38,367|
|Model 2: MLP with 1-dimensional embeddings                                    | 60,979| 37,605|
|Model 3: GLM (gamma/log link), categorical variables replaced with embeddings | 63,170| 38,360|
|Model 4: MLP with multidimensional embeddings                                 | 60,973| 37,574|
|Linear regression, predictions capped below at 0.01                           | 60,879| 38,431|

<!-- # Extending the model {#extend}

Investigate approaches for using embedding layers for numeric variables, or numeric variables that were captured as 
categorical variables (Can the Modeling of numerical variables benefit from the application of embedding layers?) How 
should this best be done? CatBoost - cut into 256 groups smoothness Map numeric to a dense layer and use that as 
embedding - basis expansion Add a distribution to the embeddings Can insights from credibility theory lead to enhanced embeddings? Take credibility mixture over embeddings Cluster embeddings PCA of embeddings
-->
# Attention based Modeling {#attention_models}
In this section we investigate whether the predictive performance of the models applied to the NFIP dataset can be improved
by applying the attention based models defined in Section \ref{attention}. We discuss the implementation of each of the 
Simple Attention, TabTransformer and TabNet models in the following sub-section, then discuss how the modifications made
to the embeddings through the application of attention can be visualized in the following sub-section. We conclude the 
section with a discussion of the model results.

@@

### Model 5: Simple Attention Network
Model 5 is conceptually similar to Model 4, except for the following two aspects. First, we standardize the dimension of the embedding 
layers to 8 dimensions for each categorical covariate. This is done to allow for the application of the attention component, which 
requires the embeddings for each covariate to have the same dimension. Second, we apply a self-attention mechanism within the network, 
as shown in Equation \ref{eq:embed_to_attn_to_nn}. We illustrate this architecture in Figure \@ref(fig:nn-2).

@@

### Model 6: TabTransformer
Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to
similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences,
Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_col = 1$. The TabTransformer
architecture is shown in Figure \@ref(fig:nn-3).

```{r nn-3, echo = FALSE, fig.cap = "Architecture of Model 6.", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/tabt.pdf")
```
### Model 7: TabNet
The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except 
that a more specialized form of self-attention is used to select the most important covariates for the Modeling problem.

## Training and Inference {#training-inference-attn}
We follow a similar training setup for the three attention based models, to the neural networks discussed in the previous section. For Models 5 and 6, 
we utilize the \texttt{Adam} optimizer as before, however, due to the increased complexity of the models in this section,
we use learning rate of 0.001 (compared to 0.01 for the simpler models). Reflecting this reduced learning rate, 
we train on the analysis set for a maximum of 150 epochs, and select the best model based on the lowest loss achieved on the assessment. Moreover, since these models are
significantly more complex than those tested in the previous section, we also add some minimal regularization using dropout [@srivastava2014dropout] to the attention layers 
and the last layer of the network, to prevent overfitting. Dropout functions by setting some components of the network randomly to zero during each epoch of training. 
This forces each component of the network to learn a somewhat independent representation of the data, thus helping to avoid overfitting by preventing components of the 
network from becoming too specialized. In this section, we drop out components with a probability of 2.5\%. A final detail is that, 
for both of these models, we use an embedding dimension of size 16 for each of the categorical covariates, since the attention mechanism requires that the embeddings 
be of a common length. For the TabTransformer model, we use a row embedding of dimension 4.

@@

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, 
as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) 
and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, outperforms Model 4 on both the RMSE and 
MAE metrics. Thus, there is some room to argue that for tabular data Modeling that uses embeddings for categorical data, 
the use of attention within models should be considered. Disappointingly, the TabNet architecture performs poorly, perhaps due to insufficient hyperparameter tuning, 
but indicating that ``out of the box", TabNet does not appear to work well for this Modeling task. Finally, the TabTransformer model shows the best performance of all
 models considered in this research, beating the other models by a significant margin on both the RMSE and MAE metrics.

Table: (\#tab:cv-results-attn) Cross validation results for Models 5-7.

|Model                                                                          |   RMSE|    MAE|
|:-----------------------------------------------------------------------------|------:|------:|
|Model 4: MLP with multidimensional embeddings                                 | 60,973| 37,574|
|Model 5: Simple Attention                                                     | 60,601| 36,739|
|Model 6: TabNet                                                               | 62,938| 38,386|
|Model 7: TabTransformer                                                       | 59,900| 36,343|
|Linear regression, predictions capped below at 0.01                           | 60,879| 38,431|

@@

Here, we explore what the TabTransformer model has learned from the data via the attention process that forms the core of the model. We focus on the difference 
between the flood embedding learned by the model, compared with the output of the Transformer layer that processes the embeddings before these enter the model. 
@Huang2020 call the latter ``contextual embeddings", because the attention mechanism within the TabTransformer model augments the embeddings based on the 
other covariates that appear for each observation $i$.

@@

```{r embdexp, echo = FALSE, fig.cap = "Embedding versus Contextual Embedding, Flood Zone, first PCA components, points coloured based on crawl space variable", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/emb_vs_cembd_exp.pdf")
```
To investigate the predictive value added by the contextual embeddings, a simple Gamma GLM was fit to these 10 000 observations, with the goal of predicting the 
claim amount based only on either the embeddings or the contextual embeddings. As shown in Table \@ref(tab:emd-vs-cembd), the performance of the latter model, 
as indicated by the RMSE and MAE metrics is significantly better, in other words, the contextual embeddings that incorporate information across all of the categorical variables are significantly better predictors 
of claim severity compared with the embeddings.

>>>>

```{=tex}
\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}
```
where the regression parameters (weights) for each layer $k \in [1;K+1]$ are represented by the matrices $a_k$ and the intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the hyperbolic tangent (tanh) function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation). Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

```{=tex}
\begin{equation} 
\label{embed_map}
  z_\mathcal{P}^j : \mathcal{P}^j \to \mathbb{R}^{q_\mathcal{P}^j }, \quad \quad p^j \mapsto  z_\mathcal{P}^j(p),
\end{equation}
```
where $q_\mathcal{P}^j$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P}^j(.)$ is an implicit function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{embed_map} states that an embedding maps a level of a categorical variable to a numerical vector. This function is left implicit, meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to specify exactly how the embeddings can be derived from the input data. In Table \@ref(tab:embed) we show an example of two dimensional embeddings for the state variables, where these have been generated randomly.

@@

```{=latex}
\begin{equation}
A: \mathbb{R}^{(q \times d) \times (q \times d)} \to \mathbb{R}^{q \times q}, ~~
(Q, K) ~\mapsto ~ A = \frac{Q~K^T}{\sqrt{d}},
\label{eq:attention_scores}
\end{equation}
```
and where the division by $\sqrt{d}$ is done element-wise. The attention scores $A$ are then normalized, or made to sum to unity, across the columns of $A$ using a so-called softmax function i.e.

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to data from 2000 to August 2019 and further subsample 100K data points to make experiments feasible on a CPU. The dataset includes a rich variety of variables, from occupancy type to flood zone. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

[^1]: <https://www.fema.gov/about/reports-and-data/openfema>

@@

-   Target variable: Amount paid on building claim

-   Predictors:

    -   $\log$ `building_insurance_coverage`
    -   `basement_enclosure_type`,
    -   `number_of_floors_in_the_insured_building`
    -   *prefix* of `flood_zone`
    -   `primary_residence`

-   Link function: $\log$

-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^2]. For example, `"A01"`, `"A02"`, and so on are recoded as simply `"A"`. A log link together with the gamma distribution is a standard choice for severity modeling, which provides a multiplicative structure where the response is positive. Following the notation in \@ref(catdata), the number of parameters in the GLM is

[^2]: <https://www.floodsmart.gov/flood-map-zone/find-yours>

```{=tex}
\begin{equation}
1 + |J^{\text{num}}| + \left(\sum_{j \in J^{\text{cat}}} n_{\mathcal{P}}^j\right) - |J^{\text{cat}}|:
(\#eq:glm1)
\end{equation}
```
one for the intercept, one for each numeric variable, and the numbers of levels minus one (due to dummy coding) for each categorical variable.

@@

| variable       | label                                       | relativity |
|:---------------|:--------------------------------------------|-----------:|
| occupancy_type | Single family residence                     |       1.00 |
| occupancy_type | 2 to 4 unit residential buidling            |       1.10 |
| occupancy_type | Non-residential building                    |       1.34 |
| occupancy_type | Residential building with more than 4 units |       1.59 |
| flood_zone     | C                                           |       1.00 |
| flood_zone     | A                                           |       1.42 |
| flood_zone     | X                                           |       1.18 |
| flood_zone     | B                                           |       1.65 |
| flood_zone     | V                                           |       1.38 |
| flood_zone     | D                                           |       1.45 |

: (\#tab:relativities) Excerpt of relativities for Model 1.

@@

| level                                       | embedding |
|:--------------------------------------------|----------:|
| 2 to 4 unit residential buidling            |      1.46 |
| Non-residential building                    |     -1.04 |
| Residential building with more than 4 units |      0.12 |
| Single family residence                     |      0.36 |

: (\#tab:embeddings-occupancy-type) Learned embeddings for `occupancy_type` in Model 2.

| level | embedding | level | embedding | level | embedding | level | embedding | level | embedding |
|:------|----------:|:------|----------:|:------|----------:|:------|----------:|:------|----------:|
| A     |      1.48 | A09   |     -0.48 | A18   |      0.39 | A28   |      1.19 | C     |      1.73 |
| A00   |     -1.05 | A0B   |      1.47 | A19   |     -1.06 | A30   |     -0.01 | D     |     -1.15 |
| A01   |     -0.18 | A10   |     -0.25 | A20   |     -1.01 | A99   |     -0.44 | V     |     -1.85 |
| A02   |     -0.06 | A11   |      0.35 | A21   |      2.08 | AE    |      1.12 | V01   |      0.02 |
| A03   |     -0.01 | A12   |     -0.38 | A22   |     -0.15 | AH    |      2.19 | V02   |     -1.55 |
| A04   |      0.05 | A13   |      2.81 | A23   |     -0.76 | AHB   |     -0.68 | V03   |     -0.98 |
| A05   |      1.30 | A14   |      0.58 | A24   |     -0.87 | AO    |      0.55 | V04   |      1.73 |
| A06   |     -0.76 | A15   |     -0.91 | A25   |      0.02 | AOB   |      0.08 | V05   |     -1.97 |
| A07   |     -0.62 | A16   |      0.79 | A26   |      0.00 | AR    |      2.93 | V06   |      0.43 |
| A08   |     -0.53 | A17   |      1.35 | A27   |      0.80 | B     |      0.33 | V07   |      1.91 |

: (\#tab:embeddings-flood-zone) Excerpt of learned embeddings for `flood_zone` in Model 2.

@@

```{=tex}
\begin{equation}
1 + |J^{\text{num}}| + |J^{\text{cat}}|.
(\#eq:glm2)
\end{equation}```
Hence, the model can be written as

@@

```{=tex}
\begin{equation}
\mu_i = \beta_0 + \sum_{j \in J^{\text{num}}}\beta_{j} x_{i,j} + \sum_{j \in J^{\text{cat}}} \beta_{j} z_{i,j},
\end{equation}
```
where $z_{i,j}$ represents the embedding value for $i$th instance of the categorical column $X_j$.

@@

| term                                     | estimate |
|:-----------------------------------------|---------:|
| (Intercept)                              |    3.371 |
| total_building_insurance_coverage        |    0.586 |
| basement_enclosure_crawlspace_type       |    0.597 |
| number_of_floors_in_the_insured_building |   -0.278 |
| occupancy_type                           |   -0.075 |
| flood_zone                               |   -0.007 |
| primary_residence                        |   -0.062 |
| community_rating_system_discount         |    0.706 |

: (\#tab:glm2) Fitted coefficients of Model 3.

@@

| level                                       |    e1 |    e2 |    e3 |
|:--------------------------------------------|------:|------:|------:|
| 2 to 4 unit residential buidling            |  0.37 | -1.76 |  1.36 |
| Non-residential building                    |  0.13 |  1.47 |  1.77 |
| Residential building with more than 4 units | -1.12 | -0.15 | -1.43 |
| Single family residence                     |  0.85 |  0.16 | -0.99 |

: (\#tab:nn2-occupancy-type) Learned embeddings for `occupancy_type` in Model 4.

| dimension | value  | dimension | value  | dimension | value  | dimension | value  |
|:----------|:-------|:----------|:-------|:----------|:-------|:----------|:-------|
| e1        | -0.167 | e11       | -0.810 | e21       | -0.300 | e31       | -0.421 |
| e2        | -0.363 | e12       | 1.380  | e22       | -0.194 | e32       | 0.594  |
| e3        | -0.893 | e13       | -0.180 | e23       | -0.320 | e33       | 1.681  |
| e4        | 0.130  | e14       | -1.612 | e24       | -0.035 | e34       | 1.668  |
| e5        | -1.588 | e15       | -1.327 | e25       | -1.128 |           |        |
| e6        | -0.123 | e16       | -0.965 | e26       | 0.168  |           |        |
| e7        | -0.259 | e17       | -1.356 | e27       | -1.022 |           |        |
| e8        | 0.324  | e18       | -0.061 | e28       | 1.011  |           |        |
| e9        | -1.430 | e19       | -0.120 | e29       | -0.843 |           |        |
| e10       | -1.598 | e20       | -0.488 | e30       | -0.743 |           |        |

: (\#tab:nn2-flood-zone) Learned embeddings for level `"A00"` of variable `flood_zone` in Model 4.

@@

```{=tex}
\begin{equation}
p_{ij} = \frac{p_{j|i} + p_{i|j}}{2n}.
\end{equation}
```
On the other hand, in the target space, t-SNE assumes the neighbor selection is driven by a Student t-distribution with one degree of freedom, i.e., a Cauchy distribution. The similarity metric for a pair of points in $\mathcal{Y}$ is then defined as the joint probability

```{=tex}
\begin{equation}
q_{ij} = \frac{(1 + \norm{y_i - y_j}^2)^{-1}}{\sum_{k \neq l} (1 + \norm{y_k - y_l}^2)^{-1}}.
\end{equation}
```
With the similarity metrics defined for each space, we can then state the objective for "closeness" for the sets of points, which is the Kullback--Leibler divergence from Q to P:

```{=tex}
\begin{equation}
D_{KL}(P || Q) = \sum_{i \neq j} p_{ij} \log \frac{p_{ij}}{q_{ij}}. (\#eq:tsne-obj)
\end{equation}
```
We now circle back to Equation \@ref(eq:pji) and discuss $\sigma_i$. The variance of the Gaussian is allowed to vary depending on where the point is in the space; specifically, it should reflect the density or sparsness of surrounding points. To train this parameter, t-SNE introduces a hyperparameter, *perplexity*, which is a scalar value set by the user, and is defined as

```{=tex}
\begin{equation}
Perp(P_i) = 2^{H(P_i)},
\end{equation}```
where $H(P_i)$ denotes the entropy of $P_i$,

```{=tex}
\begin{equation}
H(P_i) = -\sum_j p_{j|i}\log_2 p_{j|i},
\end{equation}
```
and $P_i$ is the Gaussian parameterized by $x_i$ and $\sigma_i$. During training, t-SNE looks for $P_i$ that have perpelxity corresponding to the user input. Finally, the objective as stated in Equation \@ref(eq:tsne-obj) is differentiable, and is trained via gradient descent.

@@

| Model                                                                         |   RMSE |    MAE |
|:------------------------------------------------------------------------------|-------:|-------:|
| Model 1: GLM (gamma/log link)                                                 | 65,138 | 38,367 |
| Model 2: MLP with 1-dimensional embeddings                                    | 60,979 | 37,605 |
| Model 3: GLM (gamma/log link), categorical variables replaced with embeddings | 63,170 | 38,360 |
| Model 4: MLP with multidimensional embeddings                                 | 60,973 | 37,574 |
| Linear regression, predictions capped below at 0.01                           | 60,879 | 38,431 |

: (\#tab:cv-results) Cross validation results.

```{=html}
<!-- # Extending the model {#extend}

Investigate approaches for using embedding layers for numeric variables, or numeric variables that were captured as 
categorical variables (Can the Modeling of numerical variables benefit from the application of embedding layers?) How 
should this best be done? CatBoost - cut into 256 groups smoothness Map numeric to a dense layer and use that as 
embedding - basis expansion Add a distribution to the embeddings Can insights from credibility theory lead to enhanced embeddings? Take credibility mixture over embeddings Cluster embeddings PCA of embeddings
-->
```
# Attention based Modeling {#attention_models}

In this section we investigate whether the predictive performance of the models applied to the NFIP dataset can be improved by applying the attention based models defined in Section \ref{attention}. We discuss the implementation of each of the Simple Attention, TabTransformer and TabNet models in the following sub-section, then discuss how the modifications made to the embeddings through the application of attention can be visualized in the following sub-section. We conclude the section with a discussion of the model results.

@@

### Model 5: Simple Attention Network

Model 5 is conceptually similar to Model 4, except for the following two aspects. First, we standardize the dimension of the embedding layers to 8 dimensions for each categorical covariate. This is done to allow for the application of the attention component, which requires the embeddings for each covariate to have the same dimension. Second, we apply a self-attention mechanism within the network, as shown in Equation \ref{eq:embed_to_attn_to_nn}. We illustrate this architecture in Figure \@ref(fig:nn-2).

@@

### Model 6: TabTransformer

Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences, Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_col = 1$. The TabTransformer architecture is shown in Figure \@ref(fig:nn-3).

```{r nn-3, echo = FALSE, fig.cap = "Architecture of Model 6.", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/tabt.pdf")
```

### Model 7: TabNet

The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except that a more specialized form of self-attention is used to select the most important covariates for the Modeling problem.

## Training and Inference {#training-inference-attn}

We follow a similar training setup for the three attention based models, to the neural networks discussed in the previous section. For Models 5 and 6, we utilize the \texttt{Adam} optimizer as before, however, due to the increased complexity of the models in this section, we use learning rate of 0.001 (compared to 0.01 for the simpler models). Reflecting this reduced learning rate, we train on the analysis set for a maximum of 150 epochs, and select the best model based on the lowest loss achieved on the assessment. Moreover, since these models are significantly more complex than those tested in the previous section, we also add some minimal regularization using dropout [@srivastava2014dropout] to the attention layers and the last layer of the network, to prevent overfitting. Dropout functions by setting some components of the network randomly to zero during each epoch of training. This forces each component of the network to learn a somewhat independent representation of the data, thus helping to avoid overfitting by preventing components of the network from becoming too specialized. In this section, we drop out components with a probability of 2.5%. A final detail is that, for both of these models, we use an embedding dimension of size 16 for each of the categorical covariates, since the attention mechanism requires that the embeddings be of a common length. For the TabTransformer model, we use a row embedding of dimension 4.

@@

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, outperforms Model 4 on both the RMSE and MAE metrics. Thus, there is some room to argue that for tabular data Modeling that uses embeddings for categorical data, the use of attention within models should be considered. Disappointingly, the TabNet architecture performs poorly, perhaps due to insufficient hyperparameter tuning, but indicating that \`\`out of the box", TabNet does not appear to work well for this Modeling task. Finally, the TabTransformer model shows the best performance of all models considered in this research, beating the other models by a significant margin on both the RMSE and MAE metrics.

| Model                                               |   RMSE |    MAE |
|:----------------------------------------------------|-------:|-------:|
| Model 4: MLP with multidimensional embeddings       | 60,973 | 37,574 |
| Model 5: Simple Attention                           | 60,601 | 36,739 |
| Model 6: TabNet                                     | 62,938 | 38,386 |
| Model 7: TabTransformer                             | 59,900 | 36,343 |
| Linear regression, predictions capped below at 0.01 | 60,879 | 38,431 |

: (\#tab:cv-results-attn) Cross validation results for Models 5-7.

@@

Here, we explore what the TabTransformer model has learned from the data via the attention process that forms the core of the model. We focus on the difference between the flood embedding learned by the model, compared with the output of the Transformer layer that processes the embeddings before these enter the model. @Huang2020 call the latter \`\`contextual embeddings", because the attention mechanism within the TabTransformer model augments the embeddings based on the other covariates that appear for each observation $i$.

@@

```{r embdexp, echo = FALSE, fig.cap = "Embedding versus Contextual Embedding, Flood Zone, first PCA components, points coloured based on crawl space variable", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/emb_vs_cembd_exp.pdf")
```

To investigate the predictive value added by the contextual embeddings, a simple Gamma GLM was fit to these 10 000 observations, with the goal of predicting the claim amount based only on either the embeddings or the contextual embeddings. As shown in Table \@ref(tab:emd-vs-cembd), the performance of the latter model, as indicated by the RMSE and MAE metrics is significantly better, in other words, the contextual embeddings that incorporate information across all of the categorical variables are significantly better predictors of claim severity compared with the embeddings.
